Arhangelskii's \(\alpha\)-principles and selection games

### 2019-11-06 articles/selection-alpha-games/article.tex
fix bibliography

The following characterizations were given as Definition 1 by Kocinac in
[cite Kocinac selection principles related].

@@

When \((\mc A,\mc B)\) is omitted, it is assumed that \(\mc A=\mc B\) is the
collection \(\Gamma_{X,x}\) of sequences converging to some point \(x\in X\),
as introduced by Arhangelskii in [cite Arhangelskii frequency spectrum].
Provided \(\mc A\) only contains infinite sets, it's easy to see that
\(\alpha_n(\mc A,\mc B)\) implies \(\alpha_{n+1}(\mc A,\mc B)\).

@@

Such games are well-represented in the literature; see [cite Scheepers combinatorics ramsey]
for example.
We will also consider the similarly-defined games \(G_{<2}(\mc A,\mc B)\)
(\(\plII\) chooses 0 or 1 points from each choice by \(\plI\)) and \(G_{cf}(\mc A,\mc B)\)
(\(\plII\) chooses cofinitely-many points).

@@

The similarity in nomenclature follows from the observation that every non-trivial sequence in
\(C_p(X)\) converging to the zero function \(\vec{0}\) naturally defines a corresponding
\(\gamma\)-cover in \(X\), see e.g. Theorem 4 of
[Scheepers a sequential property and covering property].

The equivalence of \(\alpha_2(\Gamma_{X,x}\Gamma_{X,x})\) and 
\(\plI\notprewin G_1(\Gamma_{X,x},\Gamma_{X,x})\) was briefly asserted by Sakai
in the introduction of [cite Sakai sequence selection properties]; the similar
equivalence of \(\alpha_4(\Gamma_{X,x}\Gamma_{X,x})\) and 
\(\plI\notprewin G_{fin}(\Gamma_{X,x},\Gamma_{X,x})\) seems to be folklore.
In fact, these relationships hold in more generality.

@@

While \(\alpha_2(\mc A,\mc B)\) involves infinite intersection and
\(G_1(\mc A,\mc B)\) involves single selections, the previous result is made
more intuitive given the following result, shown for \(\mc A=\mc B=\Gamma_{X,x}\)
by Nogura in [cite product of alpha spaces].

@@

We now demonstrate the following, in the spirit of Pawlikowskii's celebrated
result that a winning strategy for the first player in the Rothberger game
may always be improved to a winning predetermined strategy [cite pawlikowskii].

@@

Note that the assumption in Theorem \ref{pawlikowskii} that \(\mc A\)
be almost-\(\Gamma\)-like cannot be omitted. In
[todo cite Clontz k-tactics in Gruenhage game]
an example of a space \(X^*\) and point \(\infty\in X^*\) 
where \(\plI\win G_1(\mc A,\mc B)\) but
\(\plI\notprewin G_1(\mc A,\mc B)\) is given, where \(\mc A\) is
the set of open neighborhoods of \(\infty\) 
(which are all uncountable), 
and \(\mc B\) is the set \(\Gamma_{X^*,\infty}\) of sequences converging to that point.
(Note that \(G_1(\mc A,\mc B)\) is called \(Gru_{O,P}(X^*,\infty)\) in that
paper, and an equivalent game \(Gru_{K,P}(X)\) is what is directly
studied. In fact, more is shown: \(\plI\) has a winning perfect-information
strategy, but for any natural number \(k\), any strategy that only uses 
the most recent \(k\) moves of \(\plII\) and the round number
can be defeated.)

@@

This corollary generalizes e.g. Theorems 26 and 30 of [cite Scheepers 1996 Ramsey],
Theorem 5 of [cite MR2119791], and Corollary 36 of [cite Clontz dual games].

@@

%Nyikos considered these half-variants
%in [cite Nyikos 1992] for \(\mc A=\mc B=\Gamma_{X,x}\), but his arguments
%can be generalized provided \(\mc A\) is \(\Gamma\)-like.

@@

\begin{lemma}[Lemma 1.2 of (cite Nyikos 92)]
Given a family \(\{A_n:n<\omega\}\) of infinite sets, there exist infinite subsets
\(A_n'\subseteq A_n\) such that \(\{A_n':n<\omega\}\) is pairwise disjoint.
\end{lemma}

@@

Another fractional version of these \(\alpha\)-principles is given as
\(\alpha_{1.5}\) in [Nyikos 92], defined in general as follows.

@@

For this paragraph we adopt the conventional assumption that
\(\Gamma_{X,x}\) is restricted to countable sets.
Nyikos showed a consistent example
where \(\alpha_2(\Gamma_{X,x},\Gamma_{X,x})\) fails to imply
\(\alpha_{1.5}(\Gamma_{X,x},\Gamma_{X,x})\), and a consistent example where
\(\alpha_{1.5}(\Gamma_{X,x},\Gamma_{X,x})\) fails to imply
\(\alpha_{1}(\Gamma_{X,x},\Gamma_{X,x})\) [cite Nyikos 92].
On the other hand, Dow showed that \(\alpha_2(\Gamma_{X,x},\Gamma_{X,x})\)
implies \(\alpha_{1}(\Gamma_{X,x},\Gamma_{X,x})\) in the Laver model
for the Borel conjecture [cite Dow 1990]; the author conjectures
that this model (specifically, the fact that
every \(\omega\)-splitting family contains an \(\omega\)-splitting
family of size less than \(\mathfrak b\) in this model) witnesses
an affirmative answer to the following question.

@@

Noting that \(\plI\win G_1(\Gamma_X,\Gamma_X)\) if and only if
\(\plI\win G_{fin}(\Gamma_X,\Gamma_X)\) [cite Kocinac], but
the same is not true of \(G_\star(\Gamma_{X,x},\Gamma_{X,x})\)
(i.e. there are \(\alpha_4\) spaces that are not \(\alpha_2\)
[cite Shakhmatov convergence algebraic structure]),
we also ask the following.

>>>>

The following characterizations were given as Definition 1 by Kocinac in
\cite{MR2417134}.

@@

When \((\mc A,\mc B)\) is omitted, it is assumed that \(\mc A=\mc B\) is the
collection \(\Gamma_{X,x}\) of sequences converging to some point \(x\in X\),
as introduced by Arhangelskii in \cite{MR0394575}.
Provided \(\mc A\) only contains infinite sets, it's easy to see that
\(\alpha_n(\mc A,\mc B)\) implies \(\alpha_{n+1}(\mc A,\mc B)\).

@@

Such games are well-represented in the literature; see \cite{MR1378387}
for example.
We will also consider the similarly-defined games \(G_{<2}(\mc A,\mc B)\)
(\(\plII\) chooses 0 or 1 points from each choice by \(\plI\)) and \(G_{cf}(\mc A,\mc B)\)
(\(\plII\) chooses cofinitely-many points).

@@

The similarity in nomenclature follows from the observation that every non-trivial sequence in
\(C_p(X)\) converging to the zero function \(\vec{0}\) naturally defines a corresponding
\(\gamma\)-cover in \(X\), see e.g. Theorem 4 of
\cite{MR1396994}.

The equivalence of \(\alpha_2(\Gamma_{X,x}\Gamma_{X,x})\) and 
\(\plI\notprewin G_1(\Gamma_{X,x},\Gamma_{X,x})\) was briefly asserted by Sakai
in the introduction of \cite{MR2280899}; the similar
equivalence of \(\alpha_4(\Gamma_{X,x}\Gamma_{X,x})\) and 
\(\plI\notprewin G_{fin}(\Gamma_{X,x},\Gamma_{X,x})\) seems to be folklore.
In fact, these relationships hold in more generality.

@@

While \(\alpha_2(\mc A,\mc B)\) involves infinite intersection and
\(G_1(\mc A,\mc B)\) involves single selections, the previous result is made
more intuitive given the following result, shown for \(\mc A=\mc B=\Gamma_{X,x}\)
by Nogura in \cite{MR812643}.

@@

We now demonstrate the following, in the spirit of Pawlikowskii's celebrated
result that a winning strategy for the first player in the Rothberger game
may always be improved to a winning predetermined strategy \cite{MR1279482}.

@@

Note that the assumption in Theorem \ref{pawlikowskii} that \(\mc A\)
be almost-\(\Gamma\)-like cannot be omitted. In
\cite{MR3467819}
an example of a space \(X^*\) and point \(\infty\in X^*\) 
where \(\plI\win G_1(\mc A,\mc B)\) but
\(\plI\notprewin G_1(\mc A,\mc B)\) is given, where \(\mc A\) is
the set of open neighborhoods of \(\infty\) 
(which are all uncountable), 
and \(\mc B\) is the set \(\Gamma_{X^*,\infty}\) of sequences converging to that point.
(Note that \(G_1(\mc A,\mc B)\) is called \(Gru_{O,P}(X^*,\infty)\) in that
paper, and an equivalent game \(Gru_{K,P}(X)\) is what is directly
studied. In fact, more is shown: \(\plI\) has a winning perfect-information
strategy, but for any natural number \(k\), any strategy that only uses 
the most recent \(k\) moves of \(\plII\) and the round number
can be defeated.)

@@

This corollary generalizes e.g. Theorems 26 and 30 of \cite{MR1378387}
Theorem 5 of \cite{MR2119791}, and Corollary 36 of \cite{clontzDualPreprint}.

@@

@@

\begin{lemma}[Lemma 1.2 of \cite{MR1195504}]
Given a family \(\{A_n:n<\omega\}\) of infinite sets, there exist infinite subsets
\(A_n'\subseteq A_n\) such that \(\{A_n':n<\omega\}\) is pairwise disjoint.
\end{lemma}

@@

Another fractional version of these \(\alpha\)-principles is given as
\(\alpha_{1.5}\) in \cite{MR1195504}, defined in general as follows.

@@

For this paragraph we adopt the conventional assumption that
\(\Gamma_{X,x}\) is restricted to countable sets.
Nyikos showed a consistent example
where \(\alpha_2(\Gamma_{X,x},\Gamma_{X,x})\) fails to imply
\(\alpha_{1.5}(\Gamma_{X,x},\Gamma_{X,x})\), and a consistent example where
\(\alpha_{1.5}(\Gamma_{X,x},\Gamma_{X,x})\) fails to imply
\(\alpha_{1}(\Gamma_{X,x},\Gamma_{X,x})\) \cite{MR1195504}.
On the other hand, Dow showed that \(\alpha_2(\Gamma_{X,x},\Gamma_{X,x})\)
implies \(\alpha_{1}(\Gamma_{X,x},\Gamma_{X,x})\) in the Laver model
for the Borel conjecture \cite{MR975638}; the author conjectures
that this model (specifically, the fact that
every \(\omega\)-splitting family contains an \(\omega\)-splitting
family of size less than \(\mathfrak b\) in this model) witnesses
an affirmative answer to the following question.

@@

Noting that \(\plI\win G_1(\Gamma_X,\Gamma_X)\) if and only if
\(\plI\win G_{fin}(\Gamma_X,\Gamma_X)\) \cite{MR2417134}, but
the same is not true of \(G_\star(\Gamma_{X,x},\Gamma_{X,x})\)
(i.e. there are \(\alpha_4\) spaces that are not \(\alpha_2\)
\cite{shakhmatov2002convergence}),
we also ask the following.
